---
title: "ADA Data Analysis"
output:
  html_document: default
  pdf_document: default
---
### All Data is taken from the 2021 and 2020 County Health Rankings

### Stata Code for County Data
``` {r}
# import delimited "/Users/nkolesky/Desktop/ADA/revised_county_health.csv", encoding(UTF-8) 
# replace obe_21=0 if obe_21<=33
# replace obe_21=1 if obe_21>33
# drop if obe_21==.
# drop if fei_20==.
# drop if pi_20==.
# drop if aeo_20==.
# drop if uni_20==.
# drop if ieq_20==.
# save "/Users/nkolesky/Desktop/ADA/usa_county_health_dichot.dta"
```

### Stata Code for State Data
``` {r}
# import delimited "/Users/nkolesky/Desktop/state_data.csv", varnames(1) encoding(UTF-8) 
# encode county, generate(substate)
# drop county
# keep if substate ==.
# save "/Users/nkolesky/Desktop/ADA/state_data.dta"
```

### Uploading County Stata .dta file
```{r}
#library(haven)
#usa_county_health_dichot <- read_dta("Desktop/ADA/usa_county_health_dichot.dta") 
#View(usa_county_health_dichot)
```

### Uploading State Stata .dta file
```{r}
#library(haven)
#state_data <- read_dta("Desktop/ADA/state_data.dta")
#View(state_data)
```

### Uploading County Non-dichot .dta file
```{r}
#library(haven)
#county_health_data <- read_dta("Desktop/ADA/county_health_data.dta")
#View(county_health_data)
```

### Read in .csv datasets
```{r}
# County Health
county_health_data <- read.csv("~/Desktop/ADA/county_health_data.csv") #Read
View(county_health_data) #Verify

#Dichot County Health
usa_county_health_dichot <- read.csv("~/Desktop/ADA/usa_county_health_dichot.csv") #Read
View(usa_county_health_dichot) #Verify

#State Level Data
state_data <- read.csv("~/Desktop/ADA/state_data.csv") #Read
View(state_data) #Verify
```

###Load in Packages
```{r}
library(arsenal)
library(beeswarm)
library(boot)
library(dpylr)
library(forcats)
library(ggbeeswarm)
library(ggcorrplot)
library(ggplot2)
library(graphics)
library(grDevices)
library(haven)
library(htmlTable)
library(knitr)
library(methods)
library(purr)
library(readr)
library(scales)
library(stats)
library(stringr)
library(tibble)
library(tidyr)
library(tidyverse)
library(utils)
library(vipor)
```
### Logistic Regression
```{r}
obelogit <- glm(obe_21 ~ fei_20 + pi_20 + aeo_20 + uni_20 + ieq_20, data = usa_county_health_dichot, family = "binomial") # Core of the analysis

summary(obelogit) # Print to verify
confint(obelogit)
exp(cbind(OR= coef(obelogit), confint(obelogit))) # 95% CI + OR

usa_county_health_dichot$obeprob <- round(fitted(obelogit), 2) #New variable and significant figures
```

## Tables
```{r}
classificationtable <- table(usa_county_health_dichot$obe_21, usa_county_health_dichot$obeprob >.5) 
classificationtable # Print to verify

sensitivity<-(classificationtable[2,2]/(classificationtable[2,2]+classificationtable[2,1]))*100 # Test the sensitivity of the analysis
sensitivity # Print to verify

specificity<-(classificationtable[1,1]/(classificationtable[1,1]+classificationtable[1,2]))*100 # Test the specificity of the analysis
specificity # Print to verify


#Table One
county_health_data$reg <- 
    factor(county_health_data$reg, 
           levels=c(1,2,3,4,5),
           labels=c("West", 
                   "Southwest", 
                   "Midwest", 
                   "Southeast", 
                   "Northeast"))  ### Establish Labels for the table
tab1 <- tableby(reg ~ obe_21 + fei_20 + pi_20 + aeo_20 + uni_20 + ieq_20, data=county_health_data) ###Build table
summary(tab1, text=TRUE) ###Verify
```

## Plots
```{r}
# Forest Plot

datobe <- data.frame(
Index = c(1, 2, 3, 4, 5),
label = c("Food Environment Index", "Physical Inactivity (%)", "Access to Exercise Opportunities", "Uninsured (%)", "Income Inequity (80%-20%)"),                          # Establishing Labels
OR = c(.9583788, 1.0698174, .9959172, .9706844, 1.0241873),  # Odds Ratio
LL = c(.8856658, 1.0539172, .9923477, .9556166, .9205606),   # Lower Limit
UL = c(1.0364314, 1.0861376, .9994818, .9859281, 1.1402790), # Upper Limit
CI = c(".8856658, 1.0364314", "1.0539172, 1.0861376", ".9923477, .9994818", ".9556166, .9859281", ".9205606, 1.1402790")
)
datobe # Verify

plot3 <- ggplot(datobe, aes(y = Index, x = OR)) +
    geom_point(shape = 18, size = 5) +  
    geom_errorbarh(aes(xmin = LL, xmax = UL), height = 0.25) +
    geom_vline(xintercept = 1, color = "red", linetype = "dashed", cex = 1, alpha = 0.5) +
    scale_y_continuous(name = "", breaks=1:5, labels = datobe$label, trans = "reverse") +
    xlab("Odds Ratio (95% CI)") + 
    ylab(" ") + 
    theme_bw() +
    theme(panel.border = element_blank(),
          panel.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.line = element_line(colour = "black"),
          axis.text.y = element_text(size = 12, colour = "black"),
          axis.text.x.bottom = element_text(size = 12, colour = "black"),
          axis.title.x = element_text(size = 12, colour = "black")) 
plot3
```
# Correlation Plot
```{r}
obecor <- dplyr::select_if(usa_county_health_dichot, is.numeric)
obeplot <- cor(obecor, use="complete.obs")
round(obeplot, 2)        # Verify
ggcorrplot(obeplot)      # Correlation Plot
```
# Logistic Regression Models 

#Physical Inactivity
```{r}
model <- glm(obe_21 ~ pi_20, data=usa_county_health_dichot, family=binomial)

#define new data frame that contains predictor variable
newdata <- data.frame(pi_20=seq(min(usa_county_health_dichot$pi_20), max(usa_county_health_dichot$pi_20),len=500))

#use fitted model to predict values of obesity
newdata$obe_21 = predict(model, newdata, type="response")

#plot logistic regression curve
plot(obe_21 ~ pi_20, data=usa_county_health_dichot, col="steelblue")
lines(obe_21 ~ pi_20, newdata, lwd=2)
```
#Access to Exercise Opportunities
```{r}
model <- glm(obe_21 ~ aeo_20, data=usa_county_health_dichot, family=binomial)

#define new data frame that contains predictor variable
newdata <- data.frame(aeo_20=seq(min(usa_county_health_dichot$aeo_20), max(usa_county_health_dichot$aeo_20),len=500))

#use fitted model to predict values of obesity
newdata$obe_21 = predict(model, newdata, type="response")

#plot logistic regression curve
plot(obe_21 ~ aeo_20, data=usa_county_health_dichot, col="steelblue")
lines(obe_21 ~ aeo_20, newdata, lwd=2)
```

#Uninsured
```{r}
model <- glm(obe_21 ~ uni_20, data=usa_county_health_dichot, family=binomial)

#define new data frame that contains predictor variable
newdata <- data.frame(uni_20=seq(min(usa_county_health_dichot$uni_20), max(usa_county_health_dichot$uni_20),len=500))

#use fitted model to predict values of obesity
newdata$obe_21 = predict(model, newdata, type="response")

#plot logistic regression curve
plot(obe_21 ~ uni_20, data=usa_county_health_dichot, col="steelblue")
lines(obe_21 ~ uni_20, newdata, lwd=2)
```

#Cleveland Plot Charts for Significant Associations
```{r}
#Obesity Rate
ggplot(state_data, 
       aes(x=obe_21, 
           y=reorder(state, obe_21))) +
    geom_point(color="blue", 
               size = 2) +
    geom_segment(aes(x = 40, 
                     xend = obe_21, 
                     y = reorder(state, obe_21), 
                     yend = reorder(state, obe_21)),
                 color = "lightgrey") +
    labs (x = "Obesity Rate (%)",
          y = "",
          title = "Obesity Rate by State",
          subtitle = "County Health Rankings Data - 2021") +
    theme_minimal() + 
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())

#Physical Inactivity
ggplot(state_data,
      aes(x=pi_20,
          y=reorder(state, pi_20))) +
    geom_point(color="blue",
               size = 2) +
    geom_segment(aes(x = 40,
                     xend = pi_20,
                     y = reorder(state, pi_20),
                     yend = reorder(state, pi_20)),
                 color = "lightgrey") +
    labs (x = "Physical Inactivity (%)",
          y = "",
          title = "Physical Inactivity by State",
          subtitle = "County Health Rankings Data - 2020") +
    theme_minimal() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())

# Access to Exercise Opportunity
ggplot(state_data, 
       aes(x=aeo_20, 
           y=reorder(state, aeo_20))) +
    geom_point(color="blue", 
               size = 2) +
    geom_segment(aes(x = 40, 
                     xend = aeo_20, 
                     y = reorder(state, aeo_20), 
                     yend = reorder(state, aeo_20)),
                 color = "lightgrey") +
    labs (x = "Access to Exercise Opportunities (%)",
          y = "",
          title = "Access to Exercise Opportunity by State",
          subtitle = "County Health Rankings Data - 2020") +
    theme_minimal() + 
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())

#Uninsurance Rate
ggplot(state_data, 
         aes(x=uni_20, 
             y=reorder(state, uni_20))) +
      geom_point(color="blue", 
                 size = 2) +
      geom_segment(aes(x = 40, 
                       xend = uni_20, 
                       y = reorder(state, uni_20), 
                       yend = reorder(state, uni_20)),
                   color = "lightgrey") +
      labs (x = "Percent of People Uninsured (%)",
            y = "",
            title = "Uninsurance Rate by State",
            subtitle = "County Health Rankings Data - 2020") +
      theme_minimal() + 
      theme(panel.grid.major = element_blank(),
           panel.grid.minor = element_blank())
```

#Beeswarm Plots
```{r}
#Establish Values
plotdata <- county_health_data %>%
    group_by(reg) %>%
    summarize(n = n(),
        mean = mean(obe_21),
        sd = sd(obe_21),
        se = sd / sqrt(n),
        ci = qt(0.975, df = n - 1) * sd / sqrt(n))

#Plot the Regional Data
ggplot(county_health_data, 
       aes(x = factor(reg,
                      labels = c("West",
                                 "Southwest",
                                 "Midwest", "Southeast", "Northeast")), 
           y = obe_21, 
           color = reg)) +
    geom_quasirandom(alpha = 0.7,
                     size = 1.5) + 
    scale_y_continuous(label = ) +
    labs(title = "Obesity Rate by Region", 
         subtitle = "Obesity Rate for 2021",
         x = "",
         y = "Obesity Rate (%)") +
    theme_minimal() +
    theme(legend.position = "none")
```